Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study

### 2021-11-04 submission/manuscript.rmd
writing results

@@

As a secondary analysis, we evaluated treatment effect heterogeneity using a risk-based approach. The

@@

- **IBM MarketScan Medicare Supplemental Database (MDCR)** is a claims database
  representing retirees in the United States with primary or Medicare
supplemental coverage.
- **Optum De-Identified Clinformatics® Data Mart Database – Date of Death
  (Optum-DOD)** is a claims database including members of private health
insurance.
- **Optum de-identified Electronic Health Record Dataset (Optum-EHR)** is an
  EHR database including Humedica’s Electronic Health Record.

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall treatment effect
estimates of teriparatide compared to oral bisphosphonates regarding the three
outcomes of interest. To account for potential measured confounding we
developed separate large-scale propensity score models within each database
based on LASSO logistic regression using the same predefined set of measured
covariates [@Tian2018]. We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate fitted in the 1:4
propensity score-matched subset of the considered patient population. Results
from different databases were summarized using random effects meta-analysis.

@@

\newpage
# Results

@@

>>>>

options(scipen=1, digits=2)

@@

We evaluated treatment effect heterogeneity using two different
risk-based subgrouping approaches. After internally developing a hip
fracture risk model within each database, we first assessed relative
and absolute treatment effects in patients falling in the lowest 75%
risk and the highest 25% risk subgroups. As a secondary approach, we
separated patients based on current guidelines for targeting
teriparatide treatment.... (EXPAND!!!!)

@@

- **IBM MarketScan Medicare Supplemental Database (CCAE)** is a claims database
  representing retirees in the United States with primary or Medicare
supplemental coverage.
- **Optum De-Identified Clinformatics® Data Mart Database – Date of Death
  (Optum-DOD)** is a claims database including members of private health
insurance.
- **Optum de-identified Electronic Health Record Dataset (Optum-EHR)** is an
  EHR database including Humedica’s Electronic Health Record.

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall
treatment effect estimates of teriparatide compared to oral
bisphosphonates regarding the three outcomes of interest. To account
for potential measured confounding we developed separate large-scale
propensity score models within each database based on LASSO logistic
regression using the same predefined set of measured covariates
[@Tian2018]. We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate fitted in the 1:4
propensity score-matched subset of the considered patient
population. Hazard ratios (HR) derived in different databases were
summarized using random effects meta-analysis.

@@

\newpage
# Results
We extracted information on a total of 28,657 patients receiving
teriparatide and 1.02 million patients receiving an oral
bisphosphonate across the 3 databases (Table 1).

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
calibrateOverallResults <- readRDS(here::here("data/processed/metaCalibrateOverall.rds"))
```

## Overall results
In our overall analyses we found no difference between teriparatide
and oral bisphosphonates in terms of hip fracture risk
(HR 
`r calibrateOverallResults %>% filter(outcome == "101") %>% pull(hr)`;
`r calibrateOverallResults %>% filter(outcome == "101") %>% pull(lower)`
to 
`r calibrateOverallResults %>% filter(outcome == "101") %>% pull(upper)`,
95% CI). We estimated major osteoporotic fracture risk increase (HR 
`r calibrateOverallResults %>% filter(outcome == "102") %>% pull(hr)`;
`r calibrateOverallResults %>% filter(outcome == "102") %>% pull(lower)` 
to 
`r calibrateOverallResults %>% filter(outcome == "102") %>% pull(upper)`,
95% CI) and vertebral fracture risk increase (HR 
`r calibrateOverallResults %>% filter(outcome == "103") %>% pull(hr)`;
`r calibrateOverallResults %>% filter(outcome == "103") %>% pull(lower)`
to `r calibrateOverallResults %>% filter(outcome == "103") %>% pull(upper)`,
95% CI) with teriparatide (Figure \@ref(fig:plotMeta)).

## Treatment effect heterogeneity

```{r, echo=FALSE} 
metaCalibrateRiskStratified_25_75 <- readRDS(
  here::here("data/processed/metaCalibrateRiskStratified_age_50_tr_1_q_25_75.rds")
) %>%
  mutate(
    hr = d(hr),
    lower = d(lower),
    upper = d(upper)
  )
calibrateRiskStratified_25_75 <- readRDS(
  here::here("data/processed/calibrateRiskStratified_age_50_tr_1_q_25_75.rds")
) %>%
  mutate(
    hr = d(exp(logRr)),
    lower = d(exp(logLb95Rr)),
    upper = d(exp(logUb95Rr))
  )
hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 <- readRDS(
  here::here("data/processed/hipFractureAbsolute_age_50_tr_1_q_25_75_101_101.rds")
) %>%
  filter(riskStratum == "Q2") %>%
  mutate(
    estimate = d(100 * estimate),
    lower    = d(100 * lower),
    upper    = d(100 * upper)
  )

hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 <- readRDS(
  here::here("data/processed/hipFractureAbsolute_age_50_tr_1_q_25_75_101_102.rds")
) %>%
  mutate(
    estimate = d(100 * estimate),
    lower    = d(100 * lower),
    upper    = d(100 * upper)
  )

hipFractureAbsolute_age_50_tr_1_q_25_75_101_103 <- readRDS(
  here::here("data/processed/hipFractureAbsolute_age_50_tr_1_q_25_75_101_103.rds")
) %>%
  mutate(
    estimate = d(100 * estimate),
    lower    = d(100 * lower),
    upper    = d(100 * upper)
  )

hipRiskPerformance <- readRDS(here::here("data/raw/predictionPerformance.rds")) %>%
  filter(
    cohort == "EntirePopulation",
    stratOutcome == 101
  )
```

The internally developed prediction models for hip fracture showed adequate performance achieving AUC of 
`r hipRiskPerformance %>% filter(database == "ccae") %>% pull(auc) %>% d()`,
`r hipRiskPerformance %>% filter(database == "optum_extended_dod") %>% pull(auc) %>% d()`
and `r hipRiskPerformance %>% filter(database == "optum_ehr") %>% pull(auc) %>% d()`
in CCAE, Optum-DOD and Optum-EHR respectively. When assessing
heterogeneity of treatment effect for hip fracture using a 75-25% risk
split, we estimated calibrated hazard ratios of
`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "ccae") %>% pull(hr)` 
(`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "ccae") %>% pull(lower)` 
to 
`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "ccae") %>% pull(upper)`;
95% CI),
`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "optum_extended_dod") %>% pull(hr)`
(`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "optum_extended_dod") %>% pull(lower)`
to 
`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "optum_extended_dod") %>% pull(upper)`;
95% CI) and 
`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "optum_ehr") %>% pull(hr)`
(`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "optum_ehr") %>% pull(lower)`
to
`r calibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101, database == "optum_ehr") %>% pull(upper)`;
95% CI) within the highest risk quarter of CCAE, Optum-DOD and Optum-EHR respectively. This translated in absolute risk reduction of
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "ccae") %>% pull(estimate)`% (`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "ccae") %>% pull(lower)`%
to
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "ccae") %>% pull(upper)`%;
95% CI),
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "optum_extended_dod") %>% pull(estimate)`%
(`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "optum_extended_dod") %>% pull(lower)`%
to
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "optum_extended_dod") %>% pull(upper)`%;
95% CI), 
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "optum_ehr") %>% pull(estimate)`%
(`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "optum_ehr") %>% pull(lower)`%
to `r hipFractureAbsolute_age_50_tr_1_q_25_75_101_101 %>% filter(database == "optum_ehr") %>% pull(upper)`%;
95% CI) respectively (Figure \@ref(fig:plotAbsoluteHip)). The random
effects meta-analytic calibrated hazard ratio within the highest risk
quarter was
`r metaCalibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101) %>% pull(hr)`
(`r metaCalibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101) %>% pull(lower)`
to 
`r metaCalibrateRiskStratified_25_75 %>% filter(riskStratum == "Q2", stratOutcome == 101, estOutcome == 101) %>% pull(upper)`;
95% CI) across the 3 databases.

For major osteoporotic fractures we estimated 
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "ccae", riskStratum == "Q1") %>% pull(estimate)`% (`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "ccae", riskStratum == "Q1") %>% pull(lower)`%
to
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "ccae", riskStratum == "Q1") %>% pull(upper)`%;
95% CI),
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_extended_dod", riskStratum == "Q1") %>% pull(estimate)`%
(`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_extended_dod", riskStratum == "Q1") %>% pull(lower)`%
to
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_extended_dod", riskStratum == "Q1") %>% pull(upper)`%;
95% CI), 
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_ehr", riskStratum == "Q1") %>% pull(estimate)`%
(`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_ehr", riskStratum == "Q1") %>% pull(lower)`%
to `r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_ehr", riskStratum == "Q1") %>% pull(upper)`%;
95% CI) in the lower 75% hip fracture risk patients. This was reduced to 
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "ccae", riskStratum == "Q2") %>% pull(estimate)`% (`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "ccae", riskStratum == "Q1") %>% pull(lower)`%
to
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "ccae", riskStratum == "Q2") %>% pull(upper)`%;
95% CI),
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_extended_dod", riskStratum == "Q2") %>% pull(estimate)`%
(`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_extended_dod", riskStratum == "Q2") %>% pull(lower)`%
to
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_extended_dod", riskStratum == "Q2") %>% pull(upper)`%;
95% CI), 
`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_ehr", riskStratum == "Q2") %>% pull(estimate)`%
(`r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_ehr", riskStratum == "Q2") %>% pull(lower)`%
to `r hipFractureAbsolute_age_50_tr_1_q_25_75_101_102 %>% filter(database == "optum_ehr", riskStratum == "Q2") %>% pull(upper)`%;
95% CI),

For the secondary outcomes (major
osteoporotic fracture and vertebral fracture) we found no evidence of
risk-based treatment effect heterogeneity. This was anticipated as
overall effects of teriparatide on theses outcomes were not
significant at the 95% confidence level. The full set of results can
be explored at https://arekkas.shinyapps.io/ter_bis_3dbs/.

@@

\newpage
```{r plotMeta, echo=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotMeta.tiff")))
```
